Measuring contrast processing in the visual system using SSVEP (and VEP?)

### 2024-03-20 review.qmd

Neurons in the visual areas of the brain are primarily responsive to differences between light and dark in an image. This property, referred to as contrast, sets the fundamental limits of our visual abilities, which remain steady over a remarkably wide range of environmental light levels. The human response to contrast can be studied using many different techniques. Early work used psychophysical methods to measure contrast sensitivity [@Campbell1965], defined as the inverse of the lowest contrast that can be reliably detected. But neural responses can also be measured more directly using techniques such as fMRI, MEG, and EEG. Here we will describe how an EEG method known as the steady state visually evoked potential (SSVEP) technique has contributed to our understanding of human contrast processing in health, disease and throughout development.

@@

inputFrequencies_Hz=np.array([5,7])

duration_S=1

>>>>

## ABSTRACT:

Cone photoreceptor contrast modulations are the currency of the early visual system. Measuring the way that sensitivity to contrast depends on factors such as spatial and temporal frequency, age, eccentricity, chromaticity and the presence of other stimuli has been a focus of vision science for at least 100 years. One of the most productive experimental approaches in this field has been the use of the 'steady-state visually-evoked potential' (SSVEP): a technique where contrast modulating inputs are 'frequency tagged (presented at well-defined frequencies and phases) and the electrical signals that they generate in the brain are analyzed in the temporal frequency domain. SSVEPs have several advantages over conventional measures of visually-evoked responses: they have relatively unambiguous ouput measures, a high SNR and they allow us to analyze interactions between stimulus components using a convenient mathematical framework. Here we describe how SSVEPs have been used to study visual contrast over the past 70 years[@dawson_summation_1954]. Because our thinking about SSVEPs is best described in simple mathematical models, we embed code that illustrates key steps in the modelling and analysis. This document can therefore be used both as a review of the use of SSVEP in measuring human contrast processing, and as an interactive learning aid.

## INTRODUCTION:
Neurons in the visual areas of the brain are primarily responsive to differences between light and dark in an image. This property, referred to as contrast, sets the fundamental limits of our visual abilities, which remain steady over a remarkably wide range of environmental light levels. The human response to contrast can be studied using many different techniques. Early work used psychophysical methods to measure contrast sensitivity [@Campbell1965], defined as the inverse of the lowest contrast that can be reliably detected. But neural responses can also be measured more directly using techniques such as fMRI, MEG, and EEG. Here we will describe how an EEG method known as the steady state visually evoked potential (SSVEP) technique has contributed to our understanding of human contrast processing in health, disease and throughout development.

@@

inputFrequencies_Hz=np.array([5,7]) # We simulate a system with two input frequencies (perhaps contrast-reversing gratings?)

duration_S=1 # All variables in this code have an _[units] where possible
